---
title: "Markdown Example"
author: "Chris Donovan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: true
    toc_depth: 3
    number_sections: 3
    theme: readable
---

```{r setup, include=FALSE}
# Set echo = TRUE for all code chunks in the document
knitr::opts_chunk$set(echo = TRUE)

# Load packages
pacman::p_load(
  dplyr,        # various data wrangling functions
  here,         # set working directory for markdowns
  snakecase,    # change between text styles
  knitr,        # various markdown things
  sjPlot,       # regressions as html output
  stargazer,    # more regression tables
  DT,           # interactive tables
  reactable,    # more interactive tables
  gapminder,    # example dataset
  ggplot2,      # graphing
  plotly,       # interactive plots
  stringr       # functions for working with character strings
)

# I won't get into the 'here' package, but you will eventually want to figure out how to use it if you make a lot of markdowns!

# Set options
options(scipen = 999)

# Load dataset
github_url <- 'https://raw.githubusercontent.com/ChrisDonovan307/cdae6590/refs/heads/main/datasets/fsci_data.rds'
con <- gzcon(url(github_url, 'rb'))
fsci <- readRDS(con)
close(con)
```

# Introduction

Use `CTRL/CMD + Shift + k` to preview your markdown. Hit the `Visual` button, or use `CTRL/CMD + Shift + F4` to switch to visual mode, which will let you edit the formatted version in real time.

At the top of the `.Rmd` file is the YAML (Yet Another Markup Language) header. It is a human-readable data serialization language. It sets some options for your markdown and gives you a nicely formatted preamble. It is currently arranged with some of my preferred settings, but feel free to play around with this and make it your own.

We have it set here to output as html, but you can just as easily produce PDF or Word documents. Note that some outputs may come out differently (or not at all) when we render to different formats. There are a bunch of built-in themes that you can explore [here](https://www.datadreaming.org/posts/2018-04-11-r-markdown-theme-gallery/2018-04-11-r-markdown-theme-gallery.html). 

Below the YAML header in the `.Rmd` document you will find the first "code chunk". You will notice that this one does not appear in the rendered document - this is because it is the `setup` chunk, and it has the setting `include=FALSE`. We use this setup chunk to set options for chunk behavior, as well as loading packages and data and such. Note that the markdown will run in a totally separate environment, so you have to load all your data and packages within the `.Rmd` file. 

# Syntax

The `#` above makes a header. A single `#` is the largest header, and extra `#`s are smaller headers.

## Smaller Header

This header is automatically numbered because of the YAML settings and the the double `#`.

### Even Smaller Header

This is three `#`s.

#### Super Tiny Header

This is four `#`s. Note that it does not show up in the table of contents because we only asked it to keep track of the first three levels.

## More Syntax

Use a single asterisk to *make font italic*.

Use double asterisks to **make font bold**.

Note that you need a blank line between paragraphs to split up text. Starting on a new line is not enough.

To make bullet points, use `-`

-   A thing
-   Another thing
    -   Sub thing

To make numbered lists, use `1.`

1.  First thing
2.  Second thing
    -   Sub thing

To `put code in-line`, use back ticks (\`\`)

```
For multiple lines of verbatim code, use triple back ticks.
x + 1 = y
```

> To make block quotes, use `>` at the start of the line.

# Code Chunks

Here we will explore some proper code chunks. You can use `CTRL/CMD + ALT + I` to create a new chunk. After the `r` comes the chunk name. This is not required, but is convenient if we hit an error because it will tell us the name of the chunk where the error was. Otherwise, it will just say "error in chunk 14" or some such. 

If we want to run code, but not show it, we can use `echo=FALSE` in the chunk options.

```{r echo_demo, echo=FALSE}
# Remove ampersands
fsci$FSCI_region <- gsub('&', 'and', fsci$FSCI_region)

# Reduce to one variable
df <- fsci[fsci$short_label == 'Prevalence of undernourishment', ]
```

Otherwise, our code chunk will be visible. Let's show off our example regression from the FSCI paper.

```{r raw_output}
lm <- lm(normvalue ~ year + FSCI_region, data = df, weights = weight)
summary(lm)
```

This shows our code and output, but it is not very nice to look at.

# Regression Outputs

## Kable

To get a cleaner output, we can convert our regression results to a data frame, then use `knitr::kable()` to create a nice looking table.

```{r clean_output}
lm_df <- broom::tidy(lm)
knitr::kable(lm_df)
```

We can take some extra steps to get the column names capitalized and the numbers rounded:

```{r cleaner_output}
lm_df_cleaner <- lm_df %>% 
  dplyr::mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  setNames(c(snakecase::to_title_case(names(.))))
knitr::kable(lm_df_cleaner)
```

## sjPlot

For a very clean regression table with less work, try the `sjPlot` package:

```{r sjPlot}
sjPlot::tab_model(
  lm, 
  p.style = 'numeric',
  digits = 3,
  show.se = TRUE,
  robust = TRUE,
  show.reflvl = TRUE,
  dv.labels = 'Undernourishment',
  pred.labels = gsub("FSCI_region", "", names(coef(lm)))
)
```

Note that this function takes the `lm` object as an input, not a data frame. It is designed to work with regression models and provides a ton of options for displaying them. Check out the documentation [here](https://strengejacke.github.io/sjPlot/index.html). This is where you really learn how to use a package. It is written by the author, with abundant vignettes and examples.

A curious hiccup with this package is that the `show.fstat` argument does not work. If you want to see why, check out the code behind the function. You can do this either by placing the cursor on the function and hitting `F2` or by using `CTRL/CMD + left click` on the function. 

## stargazer

The `stargazer` package is quite popular in econometrics. You can find a nice tutorial [here](https://www.jakeruss.com/cheatsheets/stargazer/), or a quick paper and demo arguing why you should use it from the author [here](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf). It's a pretty nice package for easily displaying regressions in LaTeX, but I wouldn't personally recommend it for non-LaTeX applications.

Here we will show how you can put two models together in the same table.

```{r stargazer, results='asis'}
# Get another regression
df2 <- fsci[fsci$short_label == 'Access to safe water', ]
lm2 <- lm(normvalue ~ year + FSCI_region, data = df2, weights = weight)

stargazer::stargazer(
  lm,
  lm2,
  type = 'html',
  font.size = 'footnotesize',
  column.labels = c('Undernourishment', 'Safe Water'),
  dep.var.labels.include = FALSE,
  covariate.labels = gsub("FSCI_region", "", names(coef(lm)))
)
```
Note that we put the `results='asis'` option into the chunk header. This is how we can get latex to show up properly as an html markdown. 

# Interactive Tables

We've already seen how to make tables above. For static tables, `knitr::kable()` is a good choice. The `kableExtra` package is also a great extension to knitr, giving you tons of options for customization. See the [docs for examples](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf).

For interactive tables, there are a couple of different options.

## DT

The `DT` package is a classic choice for interactive tables. Note that we are setting `echo=FALSE` here, so the code chunk will not be visible.

```{r DT_table, echo=FALSE}
data(gapminder)
DT::datatable(gapminder)
```

## Reactable

My personal favorite for interactive tables is `reactable`. The [documentation](https://glin.github.io/reactable/index.html) is excellent, so check it out if you're interested.

```{r reactable_table}
reactable::reactable(
  data = gapminder,
  filterable = TRUE,
  searchable = TRUE,
  outlined = TRUE,
  bordered = TRUE,
  compact = TRUE,
  striped = TRUE,
  showPageSizeOptions = TRUE
)
```

I find the options for customization here much more intuitive than `DT`, and the documentation is much easier to use.

# Plots

## Static Plots

We haven't really covered plots, but you really just throw your code in the chunk and it will appear.

### Base Plot

Base plots with the `plot()` function are available in base R. It can do just about anything. I personally find that it works great for simple plots, but more elaborate and pretty plots take more work.

```{r base_plot}
# Filter gapminder to the year 2007 only
gapminder_2007 <- gapminder[gapminder$year == 2007, ]

# Plot gapminder data
plot(
  x = gapminder_2007$gdpPercap,
  y = gapminder_2007$lifeExp,
  col = gapminder_2007$continent,
  pch = 16,
  cex = sqrt(gapminder_2007$pop) / 10000,
  ylab = 'Life Expectancy',
  xlab = 'GDP per Capita',
  main = 'Life Expectancy against GDP per Capita (2007)'
)

# Add a legend to the plot above
legend(
  "bottomright", 
  legend = levels(gapminder_2007$continent),
  col = 1:5, 
  pch = 16, 
  title = "Continent"
)
```

### ggplot2

The `ggplot2` package is one of the biggest strengths of R in my opinion. It is an excellent package for making pretty plots easily, with tons of extensions and extra packages for applications in mapping, chord diagrams, dendrograms, animations, etc. For a nice gallery of R graphs including example code, check out the [R Graph Gallery](https://r-graph-gallery.com/).

```{r ggplot2}
# Save this plot to an object so we can use it again later
gapminder_static <- gapminder %>% 
  dplyr::filter(year == 2007) %>% 
  ggplot2::ggplot(aes(x = gdpPercap, y = lifeExp, color = continent, size = pop)) +
  geom_point() +
  theme_classic() +
  labs(
    x = 'GDP per Capita',
    y = 'Life Expectancy',
    title = 'Life Expectancy against GDP per Capita (2007)'
  )

# Show plot created above
gapminder_static
```

We can change the alignment, size, and resolution of our plot in the chunk options:

```{r plot_with_options, fig.align='center', out.width="60%", fig.cap='Caption Goes Here'}
# Show same plot from last chunk but with different settings
gapminder_static
```

## Interactive Plots

What about an interactive plot? We can use the very popular `plotly` package to do this. It is native to python, but the `plotly` R package gives us an easy way to access it. It has its own syntax, but you can also use the `ggplotly()` function to convert a ggplot object to a plotly object.

```{r plotly}
# This time we'll save the plot to an object that we can call later
gapminder_interactive <- gapminder %>%
  dplyr::filter(year == 2007) %>%
  ggplot2::ggplot(aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent,
    size = pop,
    text = paste0(
      'Country: ', country, '\n',
      'Continent: ', continent, '\n',
      'GDP per capita: $',
      stringr::str_squish(format(round(gdpPercap, 0), big.mark = ',')), '\n',
      'Life Exp: ', round(lifeExp, 1), ' years\n',
      'Population: ', format(pop, big.mark = ',')
    )
  )) +
  geom_point() +
  theme_classic() +
  labs(
    x = 'GDP per Capita',
    y = 'Life Expectancy',
    title = 'Life Expectancy against GDP per Capita (2007)'
  )

# Use ggplotly function on the plot object we made above
plotly::ggplotly(gapminder_interactive, tooltip = 'text')
```

Note that you can hover over points to see more information from out `text` field, and also move, zoom, select, and download a static image of the plot.

# Optional Resources

Great quick references:

-   Cheatsheet: [R Markdown Cheatsheet](https://rstudio.github.io/cheatsheets/rmarkdown.pdf)
-   Cheatsheet: [Data Visualization with ggplot2](https://rstudio.github.io/cheatsheets/data-visualization.pdf)
-   Example graphs with code: [R Graph Gallery](https://r-graph-gallery.com/)
-   Gallery of Colors: [Colors in R](https://r-charts.com/colors/)

If you want to dive deeper:

-   DataCamp Course: [Introduction to Data Visualization with ggplot2](https://app.datacamp.com/learn/courses/introduction-to-data-visualization-with-ggplot2)
-   DataCamp Course: [Reporting with R Markdown](https://app.datacamp.com/learn/courses/reporting-with-rmarkdown)
-   Ebook: [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/)
-   Ebook: [R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/)
-   Ebook: [ggplot2: Elegant Graphics for Data Analysis (3e)](https://ggplot2-book.org/)
-   Ebook: [Interactive Web-Based Data Visualization with R, plotly, and shiny](https://plotly-r.com/index.html)
